Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-09-29 draft.tex
completing edges
section on gradient potentials over triangulated domains

>>>>

 \mwl{First estimate, alternative description}
    We use a recursive argument.
    Suppose that $0 < m \leq M$ such that there exists $w_{\iota(m)} \in W^{1,p}(\Omega_{\iota(m)})$ with $c_{0} = u - w_{\iota(m)}$.
    In particular, $\nabla w_{\iota(m)} = \nabla u$ over $\Omega_{\iota(m)}$. 
    By assumption, $T_{m}$ shares the face $F_{m}$ with $T_{\iota(m)}$. 
    % Let $\hat T$ be the reference simplex.
    % We fix affine diffeomorphisms $\varphi : \hat T \rightarrow T_{j(m)}$ and $\vartheta : \hat T \rightarrow T_{m+1}$,
    % each of which maps the convex closure of $v_0, \dots, v_{n-1}$ onto the common face $F_{m}$.
    We define $w'_{m} := w_{\iota(m)} \circ \Xi \in W^{1,p}(T_{m})$,
    where $\Xi : T_{m} \rightarrow T_{\iota(m)}$ is the unique affine diffeomorphism that leaves $F_{m}$ invariant. 
    By construction, $w'_{m} \in W^{1,p}(T_{m})$ with 
        \trace_F w_{\iota(m)} = \trace_F w'_{m}
    %Write $w''_{m} = u - w'_{m} \in W^{1,p}(T_{m})$. Then 
    There exists $w''_{m} \in W^{1,p}(T_{m})$ such that 
        \nabla w''_{m} = \nabla u_{|T_{m+1}} - \nabla w'_{m}, 
        \trace_{F_{m}} w''_{m} = 0.
    This is true because 
        \trace_{F_{m}}( \nabla u - \nabla w'_{m} ) 
        \trace_{F_{m}} \nabla u - \trace_{F_{m}} \nabla w'_{m}
        \nabla \trace_{F_{m}} u - \nabla \trace_{F_{m}} w'_{m}
        \nabla \trace_{F_{m}} u - \nabla \trace_{F_{m}} w_{\iota(m)}
        0
    Setting $w_{m} := w_{\iota(m)}$ over $\Omega_{\iota(m)}$ and $w_{m} := w'_{m} + w''_{m}$ over $T_{m}$, 
    we find 
    \begin{gather*}
        \nabla w_{m} = \nabla w'_{m} + \nabla w''_{m} = \nabla w'_{m} + \nabla u - \nabla w'_{m} = \nabla u,
        \trace_{F_{m}} w_{m} = \trace_{F_{m}} w'_{m} = \trace_{F_{m}} w_{\iota(m)}.
    \end{gather*}
    That means that $w_{m} \in W^{1,p}(\Omega_{m})$ with $\nabla w_{m} = \nabla u$ over $\Omega_{m}$. 
    In particular, $c_{0} = u - w_{m}$. 
    Lemma~\ref{lemma:mixedbconsimplex} gives 
        \| w''_{m} \|_{L^{p}(T_{m})} 
        C_{\PF,T_{m},F_{m},p} \| \nabla u - \nabla w'_{m} \|_{L^{p}(T_{m})}.
        C_{\PF,T_{m},F_{m},p} \| \nabla u      \|_{L^{p}(T_{m})} 
        C_{\PF,T_{m},F_{m},p} \| \nabla w'_{m} \|_{L^{p}(T_{m})} 
    We also employ Lemma~\ref{lemma:volumecomparison} to find 
        \| \nabla w'_{m} \|_{L^{p}(T_{m})}
        |\det(\Jacobian \Xi  )|^{-\frac 1 p} 
        \matnorm{ \Jacobian \Xi   }
        \| \nabla w_{\iota(m)} \|_{L^{p}(T_{\iota(m)})}
        \volumeratio(\calT)^{\frac 1 p} \reflectionestimate(\calT)
        \| \nabla w_{\iota(m)} \|_{L^{p}(T_{\iota(m)})}
    Putting this all together, we use 
        \| w_{m} \|_{L^{p}(T_{m})}
        \| w'_{m} \|_{L^{p}(T_{m})}
        \| w''_{m} \|_{L^{p}(T_{m})}
        \volumeratio(\calT)^{\frac 1 p} 
        \| w_{\iota(m)} \|_{L^{p}(T_{\iota(m)})} 
        C_{\PF,T_{m},F_{m},p} 
            \| \nabla u \|_{L^{p}(T_{m})} 
            \| \nabla w'_{m} \|_{L^{p}(T_{m})} 
    and 
        \| \nabla w'_{m} \|_{L^{p}(T_{m})} 
        \volumeratio(\calT)^{\frac 1 p} \reflectionestimate(\calT)
        % C_{\PF,T_{m},F_{m},p} 
        \| \nabla w_{\iota(m)} \|_{L^{p}(T_{\iota(m)})} 
        \volumeratio(\calT)^{\frac 1 p} \reflectionestimate(\calT)
        % C_{\PF,T_{m},F_{m},p} 
        \| \nabla u \|_{L^{p}(T_{\iota(m)})} 
    The existence of $w \in W^{1,p}(\Omega)$ satisfying the recursive estimate follows.

        \begin{align*}
        \| w_{m} \|_{L^{p}(T_m)}
%         &\leq 
%         \volumeratio(\calT)^{\frac m p}
%         C_{\PF,T_{0},p} 
%         \| \nabla u \|_{L^{p}( T_{0} )}
%         +
%         \sum_{l=1}^{m} 
%         \volumeratio(\calT)^{\frac {m-l} p}
%         C_{\PF,T_{l},F_{l},p} 
%         \| \nabla u \|_{L^{p}(T_{l})} 
%         \\&\qquad 
%         +
%         \sum_{l=1}^{m} 
%         \volumeratio(\calT)^{\frac {m-l} p} \reflectionestimate(\calT)
%         C_{\PF,T_{l},F_{l},p} 
%         \| \nabla u \|_{L^{p}(T_{l-1})} 
%         \\
        &\leq 
        \volumeratio(\calT)^{\frac m p}
        C_{\PF,T_{0},p} 
        \| \nabla u \|_{L^{p}( T_{0} )}
        \\&\quad 
        +
        \sum_{l=1}^{m} 
        \volumeratio(\calT)^{\frac {m-l} p}
        C_{\PF,T_{l},F_{l},p} 
        \left( 
            \| \nabla u \|_{L^{p}(T_{l})} 
            +
            \volumeratio(\calT)^{\frac {1} p}
            \reflectionestimate(\calT)
            \| \nabla u \|_{L^{p}(T_{l-1})} 
        \right)
        .
    \end{align*}
    \begin{align*}
        \| \cye{w} \|_{L^{p}(T_{\cye{M}})}
%         &\leq 
%         \rho(\calT)^{\frac \cye{M} p}
%         C_{\PF,T_{0},p} 
%         \| \nabla u \|_{L^{p}( T_{0} )}
%         +
%         \sum_{l=1}^{\cye{M}} 
%         \rho(\calT)^{\frac {\cye{M}-l} p}
%         C_{\PF,T_{l},F_{l},p} 
%         \| \nabla u \|_{L^{p}(T_{l})} 
%         \\&\qquad 
%         +
%         \sum_{l=1}^{\cye{M}} 
%         \rho(\calT)^{\frac {\cye{M}-l} p} \xi(\calT)
%         C_{\PF,T_{l},F_{l},p} 
%         \| \nabla u \|_{L^{p}(T_{l-1})} 
%         \\
        &\leq 
        \sum_{l=1}^{\cye{M}} 
        \rho(\calT)^{\frac {\cye{M}-l} p}
        C_{\PF,T_{l},F_{l},p} 
        \left( 
            \| \nabla u \|_{L^{p}(T_{l})} 
            +
            \rho(\calT)^{\frac {1} p}
            \xi(\calT)
            \| \nabla u \|_{L^{p}(T_{l-1})} 
        \right)
        \\&\quad 
        + \rho(\calT)^{\frac{\cye{M}}{p}}
        C_{\PF,T_{0},p} 
        \| \nabla u \|_{L^{p}( T_{0} )}
        .
    \end{align*}

    Let $T \in \calT$ be any $n$-simplex. 
    Now, let $T \in \calT$ be any $n$-simplex. 
    By assumption, there exists a \cye{face} path $T_0, T_1, \dots, T_M = T$, composed of $n$-simplices \cye{from $\calT$}.

    For any $1 \leq m \leq M$ we let $F_m := T_m \cap T_{m-1}$ be a face of dimension $n-1$ \cye{shared by the $n$-simplices $T_m$ and $T_{m-1}$}. 
    \cye{Recall the notation $U_{F_m} := T_m \cup T_{m-1}$.}
    For any $1 \leq m \leq M$, \cye{recall that} $F_m := T_m \cap T_{m-1}$ \cye{is}  a face of dimension $n-1$ \cye{shared by the $n$-simplices $T_m$ and $T_{m-1}$}.

    By assumption, there exists a strong path $T_0, T_1, \dots, T_M = T$, composed of $n$-simplices.

    \todo[inline]{MV: Why a second theorem? We should explain the difference of Theorem~\ref{theorem:poincarefriedrichsestimate:gradzwei} wrt Theorem~\ref{theorem:poincarefriedrichsestimate:grad}. Which is better and when?}

    \begin{figure}[t]
\begin{center}
\begin{tikzpicture}
    \coordinate (A) at (0,0);   \coordinate (B) at (2,0); \coordinate (C) at (1,1.5);
    \coordinate (D) at (0,2.9); \coordinate (E) at (2.5,3); \coordinate (F) at (3,1.25); 
    \draw[fill=gray!50] (A) -- (B) -- (C) -- cycle; 
    \draw[fill=white  ] (A) -- (C) -- (D) -- cycle; 
    \draw[fill=white  ] (C) -- (D) -- (E) -- cycle; 
    \draw[fill=white  ] (B) -- (C) -- (F) -- cycle; 
\end{tikzpicture}
\quad
\begin{tikzpicture}
    \coordinate (A) at (0,0);   \coordinate (B) at (2,0); \coordinate (C) at (1,1.5);
    \coordinate (D) at (0,2.9); \coordinate (E) at (2.5,3); \coordinate (F) at (3,1.25); 
    \draw[fill=gray!25] (A) -- (B) -- (C) -- cycle; 
    \draw[fill=gray!50] (A) -- (C) -- (D) -- cycle; 
    \draw[fill=white  ] (C) -- (D) -- (E) -- cycle; 
    \draw[fill=white  ] (B) -- (C) -- (F) -- cycle; 
\end{tikzpicture}
\quad
\begin{tikzpicture}
    \coordinate (A) at (0,0);   \coordinate (B) at (2,0); \coordinate (C) at (1,1.5);
    \coordinate (D) at (0,2.9); \coordinate (E) at (2.5,3); \coordinate (F) at (3,1.25); 
    \draw[fill=gray!25] (A) -- (B) -- (C) -- cycle; 
    \draw[fill=gray!25] (A) -- (C) -- (D) -- cycle; 
    \draw[fill=gray!50] (C) -- (D) -- (E) -- cycle; 
    \draw[fill=white  ] (B) -- (C) -- (F) -- cycle; 
\end{tikzpicture}
\quad
\begin{tikzpicture}
    \coordinate (A) at (0,0);   \coordinate (B) at (2,0); \coordinate (C) at (1,1.5);
    \coordinate (D) at (0,2.9); \coordinate (E) at (2.5,3); \coordinate (F) at (3,1.25); 
    \draw[fill=gray!25] (A) -- (B) -- (C) -- cycle; 
    \draw[fill=gray!25] (A) -- (C) -- (D) -- cycle; 
    \draw[fill=gray!25] (C) -- (D) -- (E) -- cycle; 
    \draw[fill=gray!50] (B) -- (C) -- (F) -- cycle; 
\end{tikzpicture}
\end{center}
\caption{Exhausting a domain step by step as in Theorem~\ref{theorem:poincarefriedrichsestimate:grad} and \ref{theorem:poincarefriedrichsestimate:gradzwei}.
The dark gray triangle is the one being currently added, and the light gray triangles constitute the subdomain assembled so far.}
\label{figure:exhausting}
\end{figure}

\begin{remark}
    Theorem~\ref{theorem:poincarefriedrichsestimate:grad} depends on only a few parameters of the triangulation:
    the length of any traversal from the root simplex, the ratios of the volumes of adjacent simplices,
    the Poincar\'e--Friedrichs constants on each face patch, and the Poincar\'e--Friedrichs constant of the initial simplex. 
    The aforementioned local Poincar\'e--Friedrichs constants can generally be estimated in terms of shape-parameters. 
    If the face patches of the triangulation are convex, then better estimates are possible. 
\end{remark}

We write $\Omega_{m} \subseteq \bbR^{n}$ for the domain with $\overline\Omega_m := \cup_{j=0}^{m} T_m$, 
    which is a triangulated $n$-dimensional submanifold with boundary.
    \cye{\sout{We finally write $\Omega_m := \cup_{j=0}^{m} T_m$, which is a triangulated $n$-dimensional submanifold with boundary.}}
    \todo[inline]{MV: False, $\Omega_m$ is not necessarily a triangulated $n$-dimensional submanifold with boundary, see the 7-triangles example standing for $\Omega_m$ (then the 8th one gives $\Omega_{m+1} = \Omega$). As discussed online, we fortunately do not need here a submanifold with boundary (but we will need it and have it for shellable triangulations.) Actually, here we do not need $\Omega_m$ at all.}

    \mwl{Where to place that construction?}
 We construct the potential recursively. 
 We have seen that $c_{0} = w_{0} - u_{|\Omega_{0}}$ is a constant. 
 Assume that $1 \leq m \leq M$ such that there exists $w_{\iota(m)} \in W^{1,p}(\Omega_{\iota(m)})$ 
 such that $c_{0} = w_{\iota(m)} - u_{|\Omega_{\iota(m)}}$ is the constant above. 
 % Note that this assumption is already satisfied when $m=1$.
 Hence $\nabla w_{\iota(m)} = \nabla u$ over $\Omega_{\iota(m)}$. 
 By assumption, $T_{m}$ shares the face $F_{m}$ with $T_{{\iota(m)}}$. 
 Since $$\nabla w_{F_{m}|T_{{\iota(m)}}} = \nabla w_{\iota(m)|T_{{\iota(m)}}},$$
 there exists a constant $$c_{m} := w_{\iota(m)|T_{{\iota(m)}}} - w_{F_{m}|T_{{\iota(m)}}}.$$
 We define $w'_{F_m} := w_{F_m} + c_{m} \in W^{1,p}(U_F)$.
 By construction, $$\nabla w'_{F_m} = \nabla w_{F_m} = \nabla u_{|U_{F_m}}$$ 
 and 
 $w'_{F_{m}|T_{{\iota(m)}}} = w_{\iota(m)|T_{{\iota(m)}}}$. 
 But that already implies $w'_{F_m} = w_{|F_m} - c_{0}$. 
 Defining $w_{m} \in W^{1,p}(\Omega_m)$ by extending $w_{\iota(m)}$ with $w'_{F_m|T_m}$,
 we have $c_{0} = w_{m} - u_{|\Omega_{m}}$.

 \mwl{Closure}
 We obtain the desired estimate. 
 Repeating this argument produces $w \in L^{p}(\Omega)$ 
 such that for all $n$-simplices $T \in \calT$ we have the restriction $w_{|T} \in W^{1,p}(T)$ with $w_{|T} = u + c_{0}$.
 In particular, $w \in W^{1,p}(\Omega)$ with $\nabla w = \nabla u$.

        \color{blue}
        We study the singular values of the matrix 
        \begin{align*}
            \begin{pmatrix}
            -1                 & 0
            \\ 
            -2 \|b_z\|/\|h_z\| & 1
            \end{pmatrix}
            =
            \begin{pmatrix}
            -1             & 0
            \\ 
            -2 \tan(\beta) & 1
            \end{pmatrix}
            =
            \begin{pmatrix}
            0             & \tan(\beta)
             \\ 
            -\tan(\beta) & 0
            \end{pmatrix}
            +
            \begin{pmatrix}
            -1           & -\tan(\beta)
             \\ 
            \tan(\beta) & 1
            \end{pmatrix}
            .
        \end{align*}
        It follows that $\det(\Theta_{|K}) = 1$. 
        Its singular values satisfy 
        \begin{align*}
            ( \sigma_{\max} + \sigma_{\min} )/2
            =
            \tan(\beta)
            ,
            \qquad 
            ( \sigma_{\max} - \sigma_{\min} )/2
            =
            \sqrt{ 1 + \tan^{2}(\beta) }
            .
        \end{align*}
        We arrive at 
        \begin{align*}
            \sigma_{\max} = \sqrt{ 1 + \tan^{2}(\beta) } + \tan(\beta)
            ,
            \qquad 
            \sigma_{\min} = \sqrt{ 1 + \tan^{2}(\beta) } - \tan(\beta)
            .
        \end{align*}
\begin{align*}
    \lambda^2 = \frac{1}{2} \left( a + c \pm \sqrt{ a^2 - 2 a c + c^2 + 4b^2 } \right)
\end{align*}
\begin{align*}
    \lambda 
    &= 
    \frac{1}{\sqrt 2} \sqrt{ (1+t^2) + 1 \pm \sqrt{ (1+t^2)^2 - 2 (1+t^2)   +   1 + 4t^2 } } 
    \\
    &= 
    \frac{1}{\sqrt 2} \sqrt{ t^2 + 2 \pm \sqrt{ 1 + 2t^2 + t^4 - 2 - 2 t^2 + 1 + 4t^2 } } 
    \\
    &= 
    \frac{1}{\sqrt 2} \sqrt{ t^2 + 2 \pm \sqrt{          + t^4                 + 4t^2 } } 
    \\
    &= 
    \frac{1}{\sqrt 2} \sqrt{ t^2 + 2 \pm t \sqrt{ t^2 + 4 } } 
    .
\end{align*}
\begin{align*}
    \lambda = \frac{1}{\sqrt 2} \sqrt{ t^2 + 2 \pm t \sqrt{ t^2 + 4 } } 
\end{align*}
\begin{align*}
    \left( t^2 + 2 + t \sqrt{ t^2 + 4 } \right)
    \left( t^2 + 2 - t \sqrt{ t^2 + 4 } \right)
    =
    t^{4} + 4 t^{2} + 4
    -
    t^{2} ( t^{2} + 4 )
    =
    4
\end{align*}
\begin{align*}
    \frac{1}{\sqrt 2} \sqrt{ t^2 + 2 \pm t \sqrt{ t^2 + 4 } } 
    &=
    \sqrt{1+a^2/4} \pm a/2
    \\
    \frac 1 2 \left( t^2 + 2 \pm t \sqrt{ t^2 + 4 } \right)
    &=
    1+a^2/4
    \pm
    a \sqrt{1+a^2/4}
    + 
    a^2/4
    \\&=
    1
    +
    2a^2/4
    \pm
    a \sqrt{1+a^2/4}
    \\
    1 + \frac{t^2}{2} \pm t \sqrt{ \frac{t^2}{4} + 1 } 
    &=
    1
    +
    a^2/2
    \pm
    a \sqrt{1+a^2/4}
    .
\end{align*}
\color{black}
